---
title: "Chapter8"
author: "Zhang Chen"
date: "8/3/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(rethinking)
rstan_options(auto_write = TRUE)
```

-------

# Easy

## 8E1

(3)

## 8E2

Gibbs sampling achieves higher efficiency by using adaptive proposals, such that the proposal distribution adjusts intelligently depending on the current paramter value. The limitations of Gibbs sampling include (1) using conjugate priors and likelihood distributions, and (2) inefficiency in complex models.

## 8E3

HMC can not sample from discrete parameters. This is because HMC samples parameter values by gliding on the surface of the posterior distribution; when the parameter is discrete, this can not be achieved.

## 8E4

The effective number of samples is the number of independent samples that contain the same amount of information. This number is lower than the actual number of samples, because the steps in the chain are often autocorrelated, and therefore do not provide independent information about the posterior distribution.

## 8E5

1

-------

# Medium

## 8M1

```{r 8M1, include=FALSE}
data("rugged")
d <- rugged
d <- d[complete.cases(d$rgdppc_2000), ]
d$log_gdp <- log(d$rgdppc_2000)

d.trim <- d[c("log_gdp", "cont_africa", "rugged")]

# uniform prior on sigma
m1.unif <- map2stan(
  alist(
    log_gdp ~ dnorm(mu, sigma),
    mu <- a+bR*rugged+bA*cont_africa+bAR*rugged*cont_africa,
    a ~ dnorm(0,100), 
    bR ~ dnorm(0,10), 
    bA ~ dnorm(0,10), 
    bAR ~ dnorm(0,10), 
    sigma ~ dunif(0, 10)
  ),
  data=d.trim)

# exp prior on sigma
m1.exp <- map2stan(
  alist(
    log_gdp ~ dnorm(mu, sigma),
    mu <- a+bR*rugged+bA*cont_africa+bAR*rugged*cont_africa,
    a ~ dnorm(0,100), 
    bR ~ dnorm(0,10), 
    bA ~ dnorm(0,10), 
    bAR ~ dnorm(0,10), 
    sigma ~ dexp(1)
  ),
  data=d.trim)

```

```{r 8M1-results}
# model with uniform prior on sigma
plot(m1.unif)
pairs(m1.unif)

# model with exp prior on sigam
plot(m1.exp)
pairs(m1.exp)
```

## 8M2

```{r 8M2, include=FALSE}
m2.cauchy1 <- map2stan(
  alist(
    log_gdp ~ dnorm(mu, sigma),
    mu <- a+bR*rugged+bA*cont_africa+bAR*rugged*cont_africa,
    a ~ dnorm(0,100), 
    bR ~ dnorm(0,10), 
    bA ~ dnorm(0,10), 
    bAR ~ dnorm(0,10), 
    sigma ~ dcauchy(0, 0.5)
  ),
  data=d.trim)

m2.cauchy2 <- map2stan(
  alist(
    log_gdp ~ dnorm(mu, sigma),
    mu <- a+bR*rugged+bA*cont_africa+bAR*rugged*cont_africa,
    a ~ dnorm(0,100), 
    bR ~ dnorm(0,10), 
    bA ~ dnorm(0,10), 
    bAR ~ dnorm(0,10), 
    sigma ~ dcauchy(0, 1)
  ),
  data=d.trim)

m2.cauchy3 <- map2stan(
  alist(
    log_gdp ~ dnorm(mu, sigma),
    mu <- a+bR*rugged+bA*cont_africa+bAR*rugged*cont_africa,
    a ~ dnorm(0,100), 
    bR ~ dnorm(0,10), 
    bA ~ dnorm(0,10), 
    bAR ~ dnorm(0,10), 
    sigma ~ dcauchy(0, 1.5)
  ),
  data=d.trim)

m2.cauchy4 <- map2stan(
  alist(
    log_gdp ~ dnorm(mu, sigma),
    mu <- a+bR*rugged+bA*cont_africa+bAR*rugged*cont_africa,
    a ~ dnorm(0,100), 
    bR ~ dnorm(0,10), 
    bA ~ dnorm(0,10), 
    bAR ~ dnorm(0,10), 
    sigma ~ dcauchy(0, 2)
  ),
  data=d.trim)
```

```{r 8M2-Results, include=FALSE}
# plot cauchy distributions with different scale parameters
curve(dcauchy(x,location=0, scale=0.5), from=0, to=4, col="red", xlab="Density")
curve(dcauchy(x,location=0, scale=1), from=0, to=4, col="yellow", add=TRUE)
curve(dcauchy(x,location=0, scale=1.5), from=0, to=4, col="blue",add=TRUE)
curve(dcauchy(x,location=0, scale=2), from=0, to=4, col="green", add=TRUE)

pairs(m2.cauchy1)
pairs(m2.cauchy2)
pairs(m2.cauchy3)
pairs(m2.cauchy4)

```

## 8M3

```{r 8M3, include=FALSE}
m3.100wp <- map2stan(m2.cauchy4, warmup=100, iter=2100)
m3.500wp <- map2stan(m2.cauchy4, warmup=500, iter=2500)
m3.1000wp <- map2stan(m2.cauchy4, warmup=1000, iter=3000)
m3.2000wp <- map2stan(m2.cauchy4, warmup=2000, iter=4000)
m3.3000wp <- map2stan(m2.cauchy4, warmup=3000, iter=5000)
m3.4000wp <- map2stan(m2.cauchy4, warmup=4000, iter=6000)

```

```{r 8M3-Results}
# 100 warmup
plot(m3.100wp)

# 500 warmup
plot(m3.500wp)

# 1000 warmup
plot(m3.1000wp)

# 2000 warmup
plot(m3.2000wp)

# 3000 warmup
plot(m3.3000wp)

# 4000 warmup
plot(m3.4000wp)
```

--------

# Hard

## 8H1

```{r 8H1, include=FALSE}
mp <- map2stan(
  alist(
    a ~ dnorm(0,1),
    b ~ dcauchy(0,1)
  ),
  data=list(y=1),
  start=list(a=0,b=0),
  iter=1e4, warmup=100 , WAIC=FALSE )

```

```{r 8H1-Results}
plot(mp)
```

The trace plot for b is not healthy, and occasionally shows large spikes. This is because the Cauchy distribution has heavy tails, which means large numbers are still relatively plausible, and occasionally a very large value is sampled.

## 8H2

```{r 8H2, include=FALSE}
data("WaffleDivorce")
d <- WaffleDivorce

d$MedianAgeMarriage_s <- scale(d$MedianAgeMarriage)
d$Marriage_s <- scale(d$Marriage)

d <- d[c("Divorce", "MedianAgeMarriage_s", "Marriage_s")]

# fit model 5.1
m5.1 <- map2stan(
  alist(
    Divorce ~ dnorm(mu, sigma),
    mu <- a+bA*MedianAgeMarriage_s,
    a ~ dnorm(10, 10),
    bA ~ dnorm(0, 1),
    sigma ~ dunif(0, 10)
  ), data=d)

# fit model 5.2
m5.2 <- map2stan(
  alist(
    Divorce ~ dnorm(mu, sigma),
    mu <- a+bR*Marriage_s,
    a ~ dnorm(10, 10),
    bR ~ dnorm(0, 1),
    sigma ~ dunif(0, 10)
  ), data=d)

# fit model 5.3
m5.3 <- map2stan(
  alist(
    Divorce ~ dnorm(mu, sigma),
    mu <- a+bR*Marriage_s+bA*MedianAgeMarriage_s,
    a ~ dnorm(10, 10),
    bR ~ dnorm(0, 1),
    bA ~ dnorm(0, 1),
    sigma ~ dunif(0, 10)
  ), data=d)
```

```{r 8H2-Results}
compare(m5.1, m5.2, m5.3)
```

## 8H3

```{r 8H3, include=FALSE}
N <- 100 # number of individuals
height <- rnorm(N,10,2) # sim total height of each
leg_prop <- runif(N,0.4,0.5) # leg as proportion of height
leg_left <- leg_prop*height + # sim left leg as proportion + error
rnorm( N , 0 , 0.02 )
leg_right <- leg_prop*height + # sim right leg as proportion + error
rnorm( N , 0 , 0.02 )
# combine into data frame
d <- data.frame(height,leg_left,leg_right)

m5.8s <- map2stan(
  alist(
    height ~ dnorm( mu , sigma ) ,
    mu <- a + bl*leg_left + br*leg_right ,
    a ~ dnorm( 10 , 100 ) ,
    bl ~ dnorm( 2 , 10 ) ,
    br ~ dnorm( 2 , 10 ) ,
    sigma ~ dcauchy( 0 , 1 )
  ) , data=d, chains=4,
  start=list(a=10,bl=0,br=0,sigma=1) )

m5.8s2 <- map2stan(
  alist(
    height ~ dnorm( mu , sigma ) ,
    mu <- a + bl*leg_left + br*leg_right ,
    a ~ dnorm( 10 , 100 ) ,
    bl ~ dnorm( 2 , 10 ) ,
    br ~ dnorm( 2 , 10 ) & T[0,] ,
    sigma ~ dcauchy( 0 , 1 )
  ) ,data=d, chains=4,
  start=list(a=10,bl=0,br=0,sigma=1) )
```

```{r 8H3-Results}
plot(m5.8s)
pairs(m5.8s)

plot(m5.8s2)
pairs(m5.8s2)
```

The posterior distributions of beta parameters are now both skewed, but in mirror images. This is because the bl and br parameters will need to add up to a constant value, so that restricting br to be larger than 0 also restricts bl to be smaller than a certain value.

## 8H4

```{r 8H4}
WAIC(m5.8s)
WAIC(m5.8s2)
```

The model that restricts br to be positive has a smaller effective number of parameters, because by having this restriction the model becomes less flexible.

## 8H5

```{r 8H5}
island_pop <- c(2, 3, 4, 5, 6, 5, 4, 3, 2, 1) # populations in all islands

num_weeks <- 1e5
positions <- rep(0, num_weeks) # place holder
current <- 10 # start at an island
for (i in 1:num_weeks){
  positions[i] <- current
  
  proposal <- current + sample(c(-1, 1), size=1)
  
  if (proposal > 10){ proposal <- 1}
  if (proposal < 1){ proposal <- 10}
  
  prob_move <- island_pop[proposal]/island_pop[current]
  current <- ifelse(runif(1) < prob_move, proposal, current)
}

barplot(table(positions))
```

## 8H6

```{r 8H6}
# data: 7  waters in 10 tosses
d <- data.frame(y=7, n=10)

m.8h6 <- map(
  alist(
    y ~ dbinom(n, theta)
  ), 
  data=d, start = list(theta=d$y/d$n))

post <- extract.samples(m.8h6, n=1e4)

# MCMC
num_steps <- 1e4
post.mcmc <- rep(0, num_steps) # place holder
current <- d$y/d$n # start position

posterior <- function(theta){
  
  if (theta > 1 | theta < 0){
    p <- 0
  }else{
    prior <- ifelse(theta > 0 & theta < 1, 1, 0)
    likelihood <- dbinom(x=d$y, size=d$n, prob=theta)
    p <- prior*likelihood
  }
  return(p)
}

for (i in 1:num_steps){
  post.mcmc[i] <- current
  
  # proposal distribution
  proposal <- rnorm(mean=current, sd=0.1, n=1)
  
  # move?
  prob_move <- posterior(proposal)/posterior(current)
  
  current <- ifelse(runif(1) < prob_move, proposal, current)
}

dens(post.mcmc)
dens(post, col="blue", add=TRUE)
```


