---
title: "Chapter12"
author: "Zhang Chen"
date: "8/15/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(rethinking)
```

-----

# Easy

## 12E1

(a)

## 12E2

y ~ Binomial(1, p)

logit(p) = a + a_group[group] + b*x

a ~ Normal(0, 10)

a_group[group] ~ Normal(0, sigma_group)

b ~ Normal(0, 1)

sigma_group ~ HalfCauchy(0, 2)

## 12E3

y ~ Normal(mu, sigma)

mu = a + a_group[group] + b*x

a ~ Normal(0, 20)

a_group[group] ~ Normal(0, sigma_group)

b ~ Normal(0, 1)

sigma ~ HalfCauchy(0, 2)

sigma_group ~ HalfCauchy(0, 2)

-----

# Medium

## 12M1

```{r 12M1, include=FALSE}
data("reedfrogs")
d <- reedfrogs

d$tank <- 1:nrow(d)
d$pred <- ifelse(d$pred == "pred", 1, 0)
d$big  <- ifelse(d$size == "big", 1, 0)

# model with only varying intercepts
m12.1 <- map2stan(
  alist(
    surv ~ dbinom(density, p),
    logit(p) <- a + a_tank[tank],
    a ~ dnorm(0, 1),
    a_tank[tank] ~ dnorm(0, sigma),
    sigma ~ dcauchy(0,1)
  ),
  data=d, iter=4000, chains=4)

# model with the main effect of predation
m12.2 <- map2stan(
  alist(
    surv ~ dbinom(density, p),
    logit(p) <- a + a_tank[tank]+bp*pred,
    a ~ dnorm(0, 1),
    a_tank[tank] ~ dnorm(0, sigma),
    bp ~ dnorm(0, 1),
    sigma ~ dcauchy(0,1)
  ),
  data=d, iter=4000, chains=4)

# model with the main effect of size
m12.3 <- map2stan(
  alist(
    surv ~ dbinom(density, p),
    logit(p) <- a + a_tank[tank]+bb*big,
    a ~ dnorm(0, 1),
    a_tank[tank] ~ dnorm(0, sigma),
    bb ~ dnorm(0, 1),
    sigma ~ dcauchy(0,1)
  ),
  data=d, iter=4000, chains=4)

# model with both the main effect of size and the main effect of predation
m12.4 <- map2stan(
  alist(
    surv ~ dbinom(density, p),
    logit(p) <- a + a_tank[tank]+bb*big+bp*pred,
    a ~ dnorm(0, 1),
    a_tank[tank] ~ dnorm(0, sigma),
    c(bb, bp) ~ dnorm(0, 1),
    sigma ~ dcauchy(0,1)
  ),
  data=d, iter=4000, chains=4)

# model with both main effects and the interaction effect
m12.5 <- map2stan(
  alist(
    surv ~ dbinom(density, p),
    logit(p) <- a + a_tank[tank]+bb*big+bp*pred+bbp*big*pred,
    a ~ dnorm(0, 1),
    a_tank[tank] ~ dnorm(0, sigma),
    c(bb, bp, bbp) ~ dnorm(0, 1),
    sigma ~ dcauchy(0,1)
  ),
  data=d, iter=4000, chains=4)
```

```{r 12M1-Results}
# compare parameters across models
coeftab(m12.1, m12.2, m12.3, m12.4, m12.5)
```

Models that include the effect of predation estimate the sigma paramter (i.e., variation across tanks) to be smaller than models that do not include this effect. This is because predation explains part of the variation across tanks in survival rates, and once that is taken into account, the remaining unexplained variation across tanks becomes smaller.

## 12M2

```{r 12M2}
compare(m12.1, m12.2, m12.3, m12.4, m12.5)
```

## 12M3

```{r 12M3, include=FALSE}
# model with only varying intercepts
m12.1_cauchy <- map2stan(
  alist(
    surv ~ dbinom(density, p),
    logit(p) <- a + a_tank[tank],
    a ~ dnorm(0, 1),
    a_tank[tank] ~ dcauchy(0, sigma),
    sigma ~ dcauchy(0,1)
  ),
  data=d, iter=4000, chains=4)
```

```{r 12M3}
coeftab(m12.1, m12.1_cauchy)

# plot normal and cauchy distributions
curve(dnorm(x, mean=1.5, sd=1.5), from=-3, to=6, ylab="Probability Density")
curve(dcauchy(x, location=1.5, scale=1.5), from=-3, to=6, add=TRUE, col="red")

```

Tanks numbered 2, 4, 7, 20 and 38 have much larger intercepts when using the Cauchy prior. This is because Cauchy distributions have thicker tails than normal distributions, therefore larger deviations of tank-specific intercepts from the mean intercept of all tanks are more likely to occur under a Cauchy distribution than under a normal distribution.

## 12M4

```{r 12M4, include=FALSE}
data("chimpanzees")
d <- chimpanzees

d$recipient <- NULL
d$block_id  <- d$block

# fit the cross-classified model from the chapter
m12.6 <- map2stan(
  alist(
    pulled_left ~ dbinom(1, p),
    logit(p) <- a + a_actor[actor] + a_block[block_id] + 
      (bp + bpc*condition)*prosoc_left,
    a_actor[actor] ~ dnorm(0, sigma_actor),
    a_block[block_id] ~ dnorm(0, sigma_block),
    c(a, bp, bpc) ~ dnorm(0, 10),
    c(sigma_actor, sigma_block) ~ dcauchy(0,1)
  ),
  data=d, warmup=1000, iter=6000, chains=4, cores=3)

# fit the cross-classified model from the practice
m12.7 <- map2stan(
  alist(
    pulled_left ~ dbinom(1, p),
    logit(p) <- a_actor[actor] + a_block[block_id] + 
      (bp + bpc*condition)*prosoc_left,
    a_actor[actor] ~ dnorm(a, sigma_actor),
    a_block[block_id] ~ dnorm(b, sigma_block),
    c(a, b, bp, bpc) ~ dnorm(0, 10),
    c(sigma_actor, sigma_block) ~ dcauchy(0,1)
  ),
  data=d, warmup=1000, iter=6000, chains=4, cores=3)
```

```{r 12M4-Results}
precis(m12.6)
precis(m12.7)
```

------

# Hard

## 12H1

```{r 12H1, include=FALSE}
data("bangladesh")
d <- bangladesh

d$district_id <- as.integer(as.factor(d$district))

# fixed-effects model
m12.8 <- map2stan(
  alist(
    use.contraception ~ dbinom(1, p),
    logit(p) <- a_district[district_id],
    a_district[district_id] ~ dnorm(0, 5) 
  ),
  data=d, warmup=1000, iter=6000, chains=3, cores=3)

# varying-effects model
m12.9 <- map2stan(
  alist(
    use.contraception ~ dbinom(1, p),
    logit(p) <- a_district[district_id],
    a_district[district_id] ~ dnorm(a, sigma),
    a ~ dnorm(0, 5),
    sigma ~ dcauchy(0, 1)
  ),
  data=d, warmup=1000, iter=6000, chains=3, cores=3)

```

```{r 12H1-Results}
# get the raw proportions for each district
d.prop <- aggregate(use.contraception ~ district, data=d, mean)

# get the predictions from the fixed-effects model
post <- extract.samples(m12.8)
d.prop$fixed.pred <- logistic(apply(post$a_district, 2, mean))

# get the predictions from the varying-effects model
post <- extract.samples(m12.9)
d.prop$varying.pred <- logistic(apply(post$a_district, 2, mean))

varying.pred.mean <- logistic(mean(post$a))

# count sample size in each district
d.prop$sample <- data.frame(table(d$district_id))$Freq

# make a plot of predictions of two models
plot(0, 0, type="n", xlim=c(1,60), ylim=c(0, 1), 
     xlab="District ID", ylab = "Probability of Using Contraception")

points(d.prop$fixed.pred)
points(d.prop$varying.pred, pch=16, col=rangi2)
abline(h=varying.pred.mean, lty=2)

# magnitude of disagreement between two models as a function of sample size
d.prop$discrep <- abs(d.prop$fixed.pred-d.prop$varying.pred)
plot(discrep ~ sample, data=d.prop, 
     xlab="Sample Size in Each District", ylab = "Discrepancy Between Two Models")
```

## 12H2

```{r 12H2, include=FALSE}
data("Trolley")
d <- Trolley
d$id <- as.integer(d$id)
d$case <- NULL

# fit model without varying intercepts
m12.10 <- map2stan(
  alist(
    response ~ dordlogit(phi, cutpoints),
    phi <- bA*action+bI*intention+bC*contact,    
    c(bA, bI, bC, bAI, bCI) ~ dnorm(0, 10),
    cutpoints ~ dnorm(0, 10)
  ),
  data=d, warmup=2000, iter=6000, chains=4, cores=3,
  start=list(bA=0, bI=0, bC=0, bAI=0, bCI=0, cutpoints=c(-2,-1,0,1,2,2.5)))

# fit model with varying intercepts
m12.10 <- map2stan(
  alist(
    response ~ dordlogit(phi, cutpoints[id]),
    phi <- bA*action+bI*intention+bC*contact,    
    c(bA, bI, bC, bAI, bCI) ~ dnorm(0, 10),
    cutpoints[id] ~ dnorm(a, sigma),
    a ~ dnorm(0, 10),
    sigma ~ dcauchy(0, 1)
  ),
  data=d, warmup=2000, iter=6000, chains=4, cores=3)
```

